---
title: "Project Checkpoint 3"
author: "Natalie Cartwright"
format: html
---

## Loading in packages and datasets

```{r}
library(httr)
library(tidyjson)
library(jsonlite)
library(tidyverse)
library(glue)
library(lubridate)

cases_year <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')
```



## First acessing the API about air quality worldwide

```{r}
#https://api.api-ninjas.com/v1/airquality?city=New%20York%20City&country=United%20States

my_key = "r28T1kQCHLHvPuqrJm8YnCgtmMyQJWiKcuYyuz9u"

base_url <- "https://api.api-ninjas.com/v1/airquality?"

city <- "city=New%20York%20City"
country <- "country=United%20States"

us_url <- glue("{base_url}{city}&{country}")


us_row <- GET(us_url, add_headers("X-Api-Key" = my_key))

us_data <- fromJSON(rawToChar(us_row$content))

```

## Extracting the variable total_aqi to get the total air quality index for each country

```{r}
us_total <- us_data$overall_aqi

us_total
```

## Creating a dataset that has a city for every country in the measles dataset

```{r}
cases_year$country[cases_year$year == "2025"]

country_city_tbl <- tibble(
  country = c(
    "Algeria","Benin","Botswana","Burkina Faso","Burundi","Cameroon",
    "Central African Republic","Chad","Comoros","Congo","Côte d'Ivoire","Democratic Republic of the Congo","Equatorial Guinea", "Eswatini","Ethiopia","Gabon","Gambia","Ghana","Guinea","Guinea-Bissau","Kenya","Lesotho","Liberia","Madagascar","Malawi","Mali","Mauritania","Mozambique","Namibia","Niger","Nigeria","Rwanda","Senegal","Sierra Leone","South Africa","South Sudan","Togo","Uganda","United Republic of Tanzania","Zambia","Zimbabwe", "Antigua and Barbuda", "Argentina","Bahamas","Barbados","Belize","Bolivia (Plurinational State of)","Brazil","Canada","Chile","Colombia","Costa Rica","Cuba","Dominican Republic","Ecuador","El Salvador", "Guatemala","Guyana","Haiti","Honduras","Jamaica","Mexico","Nicaragua","Panama","Peru","Saint Lucia","Saint Vincent and the Grenadines", "Trinidad and Tobago","United States of America", "Afghanistan", "Bahrain","Djibouti","Iran (Islamic Republic of)","Iraq","Jordan", "Kuwait","Lebanon","Libya","Morocco","Oman","Pakistan","Qatar","Saudi Arabia","Somalia","Sudan","Syrian Arab Republic","Tunisia","United Arab Emirates","Yemen","Albania","Armenia","Austria","Azerbaijan","Belarus","Belgium","Bosnia and Herzegovina" ,"Bulgaria", "Croatia", "Cyprus", "Czechia","Denmark","Estonia","Finland","France","Georgia","Germany","Greece","Hungary","Ireland","Israel","Italy","Kazakhstan","Kyrgyzstan","Lithuania","Luxembourg","Montenegro","Netherlands (Kingdom of the)","Norway","Poland","Portugal","Republic of Moldova","Romania","Russian Federation", "Serbia", "Slovakia", "Slovenia","Spain","Sweden","Switzerland","Tajikistan","Türkiye","Turkmenistan","Ukraine","United Kingdom of Great Britain and Northern Ireland","Uzbekistan","Bangladesh","Bhutan","Democratic People's Republic of Korea","India","Indonesia","Maldives","Myanmar","Nepal","Sri Lanka","Thailand","Timor-Leste","Australia","Brunei Darussalam", "Cambodia","China","Cook Islands","Fiji","Japan","Kiribati","Lao People's Democratic Republic","Malaysia","Marshall Islands", "Micronesia (Federated States of)","Mongolia","Nauru","New Zealand","Niue","Palau","Philippines","Republic of Korea", "Singapore","Tuvalu","Viet Nam"
  ),
  
  city = c(
    "Algiers","Cotonou","Gaborone","Ouagadougou","Bujumbura","Douala",
    "Bangui","N'Djamena","Moroni","Brazzaville","Abidjan","Kinshasa",
    "Malabo","Mbabane","Addis Ababa","Libreville","Banjul","Accra",
    "Conakry","Bissau","Nairobi","Maseru","Monrovia","Antananarivo",
    "Lilongwe","Bamako","Nouakchott","Maputo","Windhoek","Niamey",
    "Lagos","Kigali","Dakar","Freetown","Johannesburg","Juba","Lomé",
    "Kampala","Dar es Salaam","Lusaka","Harare","St. John's","Buenos Aires","Nassau","Bridgetown","Belize City","Santa Cruz","São Paulo","Toronto","Santiago","Bogotá","San José","Havana","Santo Domingo","Guayaquil","San Salvador","Guatemala City", "Georgetown", "Port-au-Prince","Tegucigalpa","Kingston","Mexico City", "Managua", "Panama City","Lima","Castries","Kingstown","Port of Spain","New York","Kabul","Manama","Djibouti","Tehran","Baghdad","Amman","Kuwait City","Beirut","Tripoli","Casablanca","Muscat","Karachi","Doha","Riyadh","Mogadishu","Khartoum","Damascus","Tunis","Dubai","Sana'a","Tirana","Yerevan","Vienna","Baku","Minsk","Brussels","Sarajevo","Sofia","Zagreb","Nicosia","Prague","Copenhagen","Tallinn","Helsinki","Paris","Tbilisi","Berlin","Athens","Budapest","Dublin","Tel Aviv", "Rome", "Almaty","Bishkek","Vilnius","Luxembourg","Podgorica","Amsterdam","Oslo","Warsaw","Lisbon","Chișinău","Bucharest","Moscow","Belgrade","Bratislava","Ljubljana","Madrid","Stockholm","Zurich","Dushanbe","Istanbul","Ashgabat","Kyiv","London","Tashkent","Dhaka","Thimphu","Pyongyang","Mumbai","Jakarta","Malé","Yangon","Kathmandu","Colombo","Bangkok","Dili","Sydney","Bandar Seri Begawan","Phnom Penh","Shanghai","Avarua",
"Suva","Tokyo","Tarawa","Vientiane","Kuala Lumpur","Majuro","Palikir",
"Ulaanbaatar","Yaren","Auckland","Alofi","Ngerulmud","Manila","Seoul",
"Singapore","Funafuti","Ho Chi Minh City"
  ))
```

## Creating a function to extract total_aqi for each country


```{r}
get_aqi <- function(city, country){
  
  base_url <- "https://api.api-ninjas.com/v1/airquality?"
  
  city <- paste0("city=", gsub(" ", "%20", city))
  
  country <- paste0("country=", gsub(" ", "%20", country))

  my_url <- glue("{base_url}{city}&{country}")

  my_key = "r28T1kQCHLHvPuqrJm8YnCgtmMyQJWiKcuYyuz9u"
  
  my_row <- GET(my_url, add_headers("X-Api-Key" = my_key))

  my_data <- fromJSON(rawToChar(my_row$content))
  
  if (length(my_data$overall_aqi) == 0){
    
    my_total_aqi <- NA
    
  }else {
    
    my_total_aqi <- my_data$overall_aqi
    
  }
  
  return(my_total_aqi)
  
}
```

## Testing the function on Algiers, Algeria
```{r}
get_aqi("Algiers", "Algeria")
```

## Running the function over the full dataset and creating a new data frame

```{r}
first_run <- map2(.x = country_city_tbl$city, .y = country_city_tbl$country, .f = ~get_aqi(city = .x, country = .y))

aqi_df <- data.frame(total_aqi = unlist(first_run))

nrow(country_city_tbl)
nrow(aqi_df)

new <- cbind(country_city_tbl, aqi_df)
```

## Data description

This is a data frame that contains every total Air Quality Index by city and
country in the year 2025. There are 168 rows for the 168 countries in the measles
data but two cities have no data: Niamey, Niger and Seoul, Republic of Korea. 


