---
title: "Measles Global Analysis"
format:
  dashboard:
    orientation: columns
    embed-resources: true
    theme: journal
    scrolling: true
    logo: measles.png
    nav-buttons:
      - icon: github
        href: https://github.com/aryanajithnair/STAT-431
      - icon: envelope
        href: mailto:ajithnai@calpoly.edu
---

# Global Measles Cases

```{r}
#| label: packages-setup-and-load-data
#| message: false
#| warning: false
library(httr)
library(tidyjson)
library(jsonlite)
library(tidyverse)
library(stringr)
library(gt)
library(paletteer)
library(ggridges)
library(broom)
library(gganimate)
library(leaflet)
library(gifski)

cases_month <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_month.csv')
cases_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')
meta_df <- read_csv("https://raw.githubusercontent.com/zeetahhh/measles-analysis/main/mg%20copy.csv")

trend_label <- function(current, previous) {
  pct_change <- (current - previous) / previous * 100
  case_when(
    pct_change < -20 ~ "declining",
    pct_change >  20 ~ "surging",
    .default = "stable"
  )
}
```

## Column - Summary Table {width="45%"}

### Row - Table 1

::: {.card title="Table 1: Global Measles Cases by Year"}
```{r}
#| label: gt-table-rq1
# --code-- Insert table 1 of global cases by year

yearly_summary <- cases_year |>
  group_by(year) |>
  summarise(
    total_cases = sum(measles_total, na.rm = TRUE),
    n_countries = n_distinct(country),
    .groups = "drop"
  ) |>
  arrange(year) |>
  mutate(
    prev_total = lag(total_cases),
    pct_change = round((total_cases - prev_total) / prev_total * 100, 1),
    trend = trend_label(total_cases, prev_total)
  ) |>
  select(year, total_cases, n_countries, pct_change, trend)

yearly_summary |>
  select(year, total_cases, n_countries, pct_change, trend) |>
  gt() |>
  cols_label(
    year = md("**Year**"),
    total_cases = md("**Total Cases**"),
    n_countries = md("**Countries Reporting**"),
    pct_change = md("**% Change from Prior Year**"),
    trend = md("**Trend**")
  ) |>
  fmt_number(columns = total_cases, use_seps = TRUE, decimals = 0) |>
  fmt_number(columns = n_countries, use_seps = TRUE, decimals = 0) |>
  fmt_number(columns = pct_change, decimals = 1, suffix = "%") |>
  data_color(
    columns = pct_change,
    fn = scales::col_numeric(
      palette = c("#D9565C", "#FFFFFF", "#647D4B"),
      domain  = c(-100, 200)
    )
  ) |>
  tab_style(
    style = cell_text(color = "#D9565C", weight = "bold"),
    locations = cells_body(columns = trend, rows = trend == "declining")
  ) |>
  tab_style(
    style = cell_text(color = "#647D4B", weight = "bold"),
    locations = cells_body(columns = trend, rows = trend == "surging")
  ) |>
  tab_style(
    style = cell_text(color = "#848C80"),
    locations = cells_body(columns = trend, rows = trend == "stable")
  ) |>
  tab_spanner(
    label = md("**Case Counts**"),
    columns = c(total_cases, n_countries)
  ) |>
  tab_spanner(
    label = md("**Year-over-Year Change**"),
    columns = c(pct_change, trend)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_spanners()
  ) |>
  tab_options(
    table.width = pct(100),
    column_labels.font.weight = "bold",
    heading.align = "left",
    table.border.top.color = "black",
    row.striping.include_table_body = TRUE
  ) |>
  sub_missing(missing_text = "—")
```
:::

### Row - Table 2 

::: {.card title="Table 2: Detection Method Proportions" fill="false"}
```{r}
#| label: gt-table-detection
# --code-- Insert table 2 of detection method prop
methods <- list(clin_prop = "clinical", 
                epi_prop = "epi",
                lab_prop = "lab")

cases_filtered <- cases_year %>% filter(!is.na(measles_clinical),
                                     !is.na(measles_total),
                                     !is.na(measles_epi_linked),
                                     !is.na(measles_lab_confirmed))
yearly_totals <- cases_filtered %>% group_by(year) %>%
  summarise(
    total = sum(measles_total),
    clinical = sum(measles_clinical),
    epi = sum(measles_epi_linked),
    lab = sum(measles_lab_confirmed),
    .groups = "drop")

props <- function(col) {
  yearly_totals[[col]] / yearly_totals$total
}

propsbyyear <- bind_cols(
  select(yearly_totals, year),
  map_dfc(methods, props)
)

plot_long <- propsbyyear %>%
  pivot_longer(cols = c(clin_prop, epi_prop, lab_prop),
               names_to = "method",
               values_to = "proportion")

propsbyyear |>
  gt() |>
  cols_label(
    year = md("**Year**"),
    clin_prop = md("**Clinical**"),
    epi_prop = md("**Epi-Linked**"),
    lab_prop = md("**Lab Confirmed**")
  ) |>
  fmt_percent(columns = c(clin_prop, epi_prop, lab_prop), decimals = 1) |>
  tab_spanner(
    label = md("**Proportion of Total Cases by Detection Method**"),
    columns = c(clin_prop, epi_prop, lab_prop)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_spanners()
  ) |>
  tab_options(
    table.width = pct(100),
    column_labels.font.weight = "bold",
    heading.align = "left",
    table.border.top.color = "black",
    row.striping.include_table_body = TRUE
  )
```
:::

## Column - Value Boxes & Chart {width="55%"}

### Row - RQ1 Value Boxes {height="165px"}
```{r}
yearly_totals  <- cases_year |>
   group_by(year) |>
   summarise(total = sum(measles_total, na.rm = TRUE))

 highest_year   <- yearly_totals |> slice_max(total, n = 1)
 lowest_year    <- yearly_totals |> slice_min(total, n = 1)
 max_pct_change <- yearly_totals |>
   mutate(pct_change = (total - lag(total)) / lag(total) * 100) |>
   slice_max(abs(pct_change), n = 1) |>
   pull(pct_change) %>% round(digits= 2)
```

::: {.valuebox icon="arrow-up-circle" color="danger"}
**Highest single-year cases**

`{r} highest_year$year`

`{r} str_c(format(highest_year$total, scientific = FALSE), "cases", sep = " ")`
:::

::: {.valuebox icon="arrow-down-circle" color="success"}
**Lowest single-year cases**

`{r} lowest_year$year`

`{r} str_c(format(lowest_year$total, scientific = FALSE), "cases", sep = " ")`
:::

::: {.valuebox icon="percent" color="info"}
**Largest year-over-year % change**

`{r} str_c(max_pct_change, "%")`
:::

### Row - Charts
::: {.panel-tabset}
## Global Trend {height="50%"}

::: {.card title="Figure 1: Global Measles Cases Over Time"}
```{r}
#| label: rq1-trend-chart
# --code-- lollipop figure 1: Global Measles Cases Over Time
ggplot(yearly_summary, aes(x = year, xend = year, y = 0, yend = pct_change,
                           color = trend)) +
  geom_segment(size = 1.3, alpha = 0.9) +
  geom_point(aes(y = pct_change)) +
  theme_bw() +
  scale_x_continuous(breaks = yearly_summary$year) +
  scale_y_continuous(breaks = seq(-100, 200, by = 50)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
    labs(title = "Global Measles Cases Over Time", subtitle = "Percent Change Year-to-Year (2012-2025)") +
  xlab("Year") + ylab("Percent Change (%)") +
  scale_color_manual(values = c(surging = "#647D4BFF",
                      declining = "#D9565CFF",
                      stable = "#848C80"))
```
:::

## Detection Method {height="38%"}

::: {.card title="Figure 2: Detection Method by Year"}
```{r}
#| label: rq1-detection-chart
# --code-- stacked bar charts Figure 2: Detection Method by Year
ggplot(plot_long, aes(x = year, y = proportion, fill = method)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_x_continuous(breaks = plot_long$year) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10)) +
  labs(title = "Measles Detection Method by Year",
       subtitle = "Clinical vs. Epi-linked vs. Lab Cases") +
  xlab("Year") +
  ylab("Proportion") +
  scale_fill_paletteer_d("nationalparkcolors::Acadia", 
                         labels = c(clin_prop = "Clinical",
                                 epi_prop = "Epi-Linked",
                                 lab_prop = "Lab Confirmed"))
```
:::

:::

### Row - Animated Global Trend {height="50%"}

::: {.card title="Figure 3: Global Measles Cases Over Time (Animated)"}
```{r}
#| warning: false
#| message: false
#| label: mad-animated
#| fig.show: 'animate'
ggplot(data = yearly_summary, aes(x = year, y = total_cases / 1000, group = 1)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(breaks = seq(50, 550, by = 100)) +
  scale_x_continuous(breaks = seq(2012, 2025, by = 2)) +
  labs(title = "Total Measles Cases by Year (Globally)") +
  xlab("Year (2012 - 2025)") + ylab("Number of Cases (in Thousands)") +
  transition_reveal(year) +
  theme_bw()
```
:::


# Madagascar Data

## Column - Country Context {width="30%"} 

::: {.card title="Why Madagascar?"}
Madagascar is an intriguing statistical outlier in the global measles data set. In 2012–2017, there were fewer than 100 cases per year (stable), but in 2018–2019 cases spiked to over **213,000** — nearly a <mark>1,700% increase</mark>. Following this peak, cases declined dramatically to fewer than 100 cases.
:::

## Column - Madagascar Visualizations {width="70%"}

### Row - Figure 4- Animated Map {height="50%"}
::: {.card title="Figure 4: Madagascar Measles Cases Over Time"}
```{r}
#| label: mad-plotly
#| warning: false
#| message: false
madagascar <- cases_year %>% filter(country == "Madagascar")

ggplot(data = madagascar, aes(x = year, y = total_population / 1000000, group = 1)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(breaks = seq(20, 35, by = 5)) +
  scale_x_continuous(breaks = seq(2012, 2025, by = 2)) +
  labs(title = "Madagascar's Population by Year") +
  xlab("Year (2012 - 2025)") +
  ylab("Population (in Millions)") +
  transition_reveal(year) +
  theme_bw()
```
:::


### Row - Interactive Maps {height="50%"}

::: {.panel-tabset}

## All Cities {height="50%"}

::: {.card title="Figure 5: Madagascar City Populations"}
```{r}
#| warning: false
#| message: false
madagascarLat <- -18.77
madagascarLon <- 46.85
madagascar_cities <- meta_df

pal <- colorNumeric(palette = "plasma", domain = madagascar_cities$population)

leaflet() %>%
  setView(lat = madagascarLat, lng = madagascarLon, zoom = 5) %>%
  addProviderTiles("Esri.WorldStreetMap") %>%
  addCircleMarkers(
    data = madagascar_cities,
    lat = ~lat,
    lng = ~lng,
    radius = 6,
    fillColor = ~pal(population),
    fillOpacity = 0.8,
    color = "white",
    weight = 1,
    popup = ~paste0("<b>", city, "</b><br>Population: ", scales::comma(population))
  ) %>%
  addLegend(
    pal = pal,
    values = madagascar_cities$population,
    title = "Population",
    position = "bottomright"
  )
```
:::


## Excluding Capital

::: {.card title="Figure 6: Madagascar City Populations (Excluding Antananarivo)"}
```{r}
#| warning: false
#| message: false
madagascarWithoutAnatananarivo <- madagascar_cities |>
  filter(city != "Antananarivo")

pal2 <- colorNumeric(palette = "plasma", domain = madagascarWithoutAnatananarivo$population)

leaflet() %>%
  setView(lat = madagascarLat, lng = madagascarLon, zoom = 5) %>%
  addProviderTiles("Esri.WorldStreetMap") %>%
  addCircleMarkers(
    data = madagascarWithoutAnatananarivo,
    lat = ~lat,
    lng = ~lng,
    radius = 6,
    fillColor = ~pal2(population),
    fillOpacity = 0.8,
    color = "white",
    weight = 1,
    popup = ~paste0("<b>", city, "</b><br>Population: ", scales::comma(population))
  ) %>%
  addLegend(
    pal = pal2,
    values = madagascarWithoutAnatananarivo$population,
    title = "Population",
    position = "bottomright"
  )
```
:::

:::